---
title: "IceTest"
author: "jromine-USFWS"
date: "`r Sys.Date()`"
output: html_document
---
*This forecast is generated from publicly available data on [PTAGIS.org](https://www.ptagis.org), [Columbia Basin Research DART](http://www.cbr.washington.edu/dart/query/pitadult_hrt), and from [Mid-Columbia Fish & Wildlife Conservation Office](https://www.fws.gov/office/mid-columbia-fish-and-wildlife-conservation)*

Passive integrated transponder (PIT) tags are used throughout the Columbia River basin to track migratory fish behavior. PIT tags are used to monitor post-release performance metrics including run timing, survival, and forecasting salmon returns. Based on tagging rates (Table 1) reported by Leavenworth National Fish Hatchery (LNFH) we can expand the PIT tag count at Bonneville Dam to forecast returns to Icicle Creek up to four weeks in advance of their arrival. This advanced notice is used to inform harvest decisions and broodstock needs for LNFH. The forecast follows the model $N = (B * C) / P$, where $B$ is the PIT count at Bonneville Dam $C$ is the survival rate between Bonneville Dam and Icicle Creek mouth (conversion rate), and $P$ is the proportion of run completed based on average run timing characteristics (Columbia Basin DART, 2018). The conversion rate is estimated using PIT tag interrogation data at the antenna array at lower Icicle Creek (ICL) and Bonneville Dam (BON). 

```{r setup, include=FALSE, warning=FALSE}

#functions
# ipak <- function(pkg){
#   new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
#   if (length(new.pkg))
#     install.packages(new.pkg, dependencies = TRUE)
#   sapply(pkg, require, character.only = TRUE)
# }#libs
library(httr)
library(readr)
library(tidyverse)
library(tibble)
library(here)
library(lubridate)
library(stringr)
library(ggplot2)
library(knitr)
library(kableExtra)
#libs=c('httr', 'readr', 'tidyverse','tibble', 'here', 'lubridate', 'stringr', 'ggplot2', 'knitr', 'kableExtra')
#ipak(libs)


#Connect Web API PTAGIS
res<-GET("https://api.ptagis.org/reporting/reports/Flora1792/file/Icicle_Creek_Forecast.csv",accept('text/plain'))#gets the API connection
res #shows you what the details of the file are
stringi::stri_enc_detect(content(res, "raw"))#tells you what content is encoded as

#Read in PTAGIS returns
returns<-content(res, encoding = "UTF-16LE")#gets you the return data in table format

#Reformat date columns as dates
returns$`Release Date` <- as.Date(returns$`Release Date`, format = "%M/%d/%Y")
returns$`First Obs Date Max` <- as.Date(returns$`First Obs Date Max`, format = "%M/%d/%Y")
returns$`Last Obs Date Min` <- as.Date(returns$`Last Obs Date Min`, format = "%M/%d/%Y")

#Make a new table with Age added 
my_returns <- select(returns,'Tag', 'Site', 'First Obs Date Max', 'Brood Year', 'Release Date') %>% 
  mutate(Age = as.integer(format(returns$`First Obs Date Max`,"%Y")) - `Brood Year`)

#Clean up column names
colnames(my_returns) <- c("Tagcode", "SiteName", "MigYear", "BroodYear", "RelYear", "Age")

#Abbreviate array name
my_returns$SiteName = substr(my_returns$SiteName, 1,3)

#Create new adult table by filtering fish from age 2-6
adult_migrants <- my_returns %>%
  filter( Age > 2 & Age < 6)

#Convert detection date to release year in first date and rel year columns
adult_migrants$MigYear <- year(adult_migrants$MigYear)
adult_migrants$RelYear <- year(adult_migrants$RelYear)

#Remove duplicate tag codes at same site, count how many tags/year at each site, 
#reformat table to show counts at each array for each year, arrange by year, and add a grand total tags 
pivot <- adult_migrants %>%
  filter(duplicated(paste(adult_migrants$Tagcode, adult_migrants$SiteName) != TRUE)) %>% # remove dupes
  count(MigYear, SiteName) %>% 
  spread(SiteName, n, fill=0) %>% 
  arrange(MigYear) %>% 
  filter(MigYear > 2011 & MigYear < year(Sys.Date())) %>%
  mutate(GrandTotal=BON+ICL) %>%
  mutate(ConversionRate = (ICL / 0.87) / BON) #%>%
  #rename(Year = MigYear)

#Icicle array efficiency, previously calculated
ICLefficiency <- 0.87

#Calculate conversion rate for BON, MCN, and ICL for years past 2011(i.e., how many fish make it past each array)
ConversionTbl <- pivot %>%
  select(MigYear,BON, MCN, ICL) %>%
  mutate("Conversion" = round( ( (ICL / ICLefficiency) / BON),3)) %>%
  filter( MigYear > 2011 & MigYear < year(Sys.Date())) 

#Find average conversion rate over the years, round to 3 dec places
conversion_rate <- round(mean(ConversionTbl$Conversion), 3)

#Read in PIT tag ratio data and rename columns

PITratios <- read.csv("PITratios.csv")

pit_ratios <- PITratios %>%
  mutate(BroodYear = Release.Year-2) %>%
  select(Release.Year, Release.Number, X..PIT, PIT.Ratio.Non.Tag.Tag, BroodYear)%>%
  rename(RelYear = Release.Year, RelNumber = Release.Number, NumberTagged = X..PIT, Ratio = PIT.Ratio.Non.Tag.Tag)

#Redefine column types, remove ,'s in number columns
pit_ratios$BroodYear <- as.integer(pit_ratios$BroodYear)
pit_ratios$RelNumber <- as.integer( gsub(",", "", pit_ratios$RelNumber))
pit_ratios$NumberTagged <- as.integer( gsub(",","", pit_ratios$NumberTagged))
```

```{r eval=FALSE, warning=FALSE, include=FALSE}

#Get current fish counts over BON
current_BONcounts <- adult_migrants %>%
  count(MigYear, BroodYear, Age, SiteName) %>%
  spread(SiteName, n, fill=0) %>%
  select(MigYear, BroodYear, Age, BON) %>%
  filter(MigYear == year(Sys.Date())) ## This years fish

#Make a data frame of ages we are interested in
current_year <- data.frame(Age = as.integer(c(3,4,5)))

#Make a new table
current_year <- current_year %>% #Use data frame of ages
  mutate(BroodYear = as.integer(year(Sys.Date()))-Age) %>% #Add a column for Brood Year
  left_join(select(current_BONcounts, BroodYear, BON), by = "BroodYear" ) #Join the current BON counts table
current_year_BON_counts<-current_year %>%
  left_join(pit_ratios, by = "BroodYear") %>% #Join the PIT tag table
  mutate("Expanded Migration Size" = BON * Ratio) %>% #Expand the migration size by PIT tag ratio
  rename("Brood Year" = BroodYear, "Bonneville Count" = BON, "Number Released" = RelNumber, "Number Tagged" = NumberTagged, "Ratio Released:Tagged" = Ratio)%>% #Clean up col names
  replace_na(list("Bonneville Count" = 0, "Expanded Migration Size" = 0)) #Replace NA values with 0

#Read in previous return data
dart<-read.csv("hrt_pitadult.csv",header = TRUE)
dart <- dart[,1:10] #keeps columns 1-10

if (format(Sys.Date(), "%j") < format(as.Date(dart[2,3], format="%d-%b"), "%j")) {
  percent_run <- 0
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,3], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,4], format="%d-%b"), "%j")){
    percent_run <- 0.05
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,4], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,5], format="%d-%b"), "%j")){
    percent_run <- 0.10
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,5], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,6], format="%d-%b"), "%j")) {
    percent_run <- 0.25
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,6], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,7], format="%d-%b"), "%j")) {
    percent_run <- 0.5
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,7], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,8], format="%d-%b"), "%j")) {
    percent_run <- 0.75
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,8], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,9], format="%d-%b"), "%j")) {
    percent_run <- 0.90
  } else if (format(Sys.Date(), "%j") > format(as.Date(dart[2,9], format="%d-%b"), "%j") &
         format(Sys.Date(), "%j") <= format(as.Date(dart[2,10], format="%d-%b"), "%j")) {
    percent_run <- 0.95
  } else percent_run <- 1

#good to here
```